TensorAlloy: an automatic atomistic neural network program for alloys

### 2019-07-30 doc/papers/nn/manuscript.tex
The manuscript is updated.

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can provide density functional theory level predictions 
within negligible time. Currently, the symmetry function descriptor based 
atomistic neural network is the most widely used model for modeling alloys. 
To precisely describe complex potential energy surfaces, integrating advanced 
metrics, such as force or virial stress, into training can be of great help. 
The traditional way of implementing machine learning models starts from 
pre-computed descriptor and the analytical (model-derived) atomic forces and 
virial stress are handled by separate codes which require plenty of efforts. 
In this work, we choose a new route to overcome this challenge by building 
computation graph from atomic positions directly. Thus, TensorFlow or other 
modern machine learning platforms can handle the derivations of forces and 
stress, both automatically and efficiently. We demonstrate our core algorithm 
\textemdash the virtual atom approach \textemdash using Behler-Parinello's 
symmetry function atomic descriptor. This algorithm is implemented in our Python 
program TensorAlloy. TensorAlloy supports constructing machine learning 
interaction potentials for both \textit{molecules} and \textit{solids}. The QM7 
and SNAP/Ni-Mo datasets are used to discuss performances of our program.
\end{abstract}

@@

In this work, we propose a new algorithm, named the virtual atom approach, to 
implement the symmetry function descriptor and the ANN model based on 
TensorFlow. This algorithm is very suitable for parallel execution. With this 
approach, stochastic gradient descent (SGD) based mini-batch training can be 
adopted. We also derive a general and vectorizable expression for computing 
virial stress for arbitrary ANN model. All these algorithms have been 
implemented in our Python program: TensorAlloy.

This paper is organized as follows. Section \ref{section:theory} introduces the 
theory of the symmetry function descriptor and the ANN framework. Section 
\ref{section:method} describes the virtual atom approach and a modified ANN 
model used by TensorAlloy. Section \ref{section:discussions} displays 
performances of TensorAlloy on two public datasets. The appendix 
\ref{section:appendix} gives derivation and implementation details of key 
algorithms. Implementations of the core algorithms and some trained models can 
be obtained from GitHub (https://github.com/Bismarrck/vap) freely.

@@

There are many choices of $F_G$. The symmetry function descriptor, proposed 
by Parinello and Behler in 2007\cite{Behler}, is currently the most widely used 
atomic descriptor. The symmetry function descriptor consists of two sets of 
invariant (tranlational, rotational and permutational) functions: the radial 
symmetry function $\mathbf{G}^{(2)}$ and the angular symmetry function 
$\mathbf{G}^{(4)}$:
\begin{align}
\label{eq:G2}
\mathbf{G}^{(2)}_i(\eta, \omega) = & \sum_{j \neq i}{
    \sum_{\mathbf{n}}{g_2 (\eta, i, j, \omega)}
} \\
\label{eq:G4}
\mathbf{G}^{(4)}_i(\beta, \gamma, \zeta) 
= & 2^{1-\zeta}\sum_{j, k\neq j, k\neq i}{
    \sum_{\mathbf{n}_1}{\sum_{\mathbf{n}_2}{\sum_{\mathbf{n}_3}{
        g_4(\beta, \gamma, \zeta, 
          i, j, k, 
          \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3)
    }}}
} \\
\label{eq:g2}
g_2(\eta, i, j, \omega) = 
& \exp\left\{-\frac{\eta(\rijn - \omega)^2}{r_c^2} \right\} \cdot f_c(\rijn) \\
\label{eq:g4}
g_4(
    \beta, \gamma, \zeta, 
    i, j, k, 
    \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3) = & 
\left( 1 + \gamma\cos\theta \right)^\zeta 
    \exp \left\{ -\frac{\beta(\rijna^2 + \rikna^2 + \rjkna^2)}{r_c^2} \right\}
\nonumber \\
& \times f_c(\rijna)f_c(\rikna)f_c(\rjkna)
\end{align}
where $\eta$, $\omega$, $\beta$, $\gamma$ and $\zeta$ are empirically chosen 
parameters, $f_c(r)$ is a cutoff (damping) function and its original form is:
\begin{equation}
\label{eq:fc_origin}
f_c(r) =
    \begin{cases}
    0 & \text{$r > r_c$} \\
    \frac{1}{2} + \frac{1}{2}\cos\left(r / r_c \cdot \pi \right) & 
    \text{$r \leq r_c$}
    \end{cases}
\end{equation}
and $r_c$ is the cutoff radius. Equation \eqref{eq:fc_origin} can be further 
transformed to a vectorized expression:
\begin{equation}
\label{eq:fc}
f_c(r) = \frac{1}{2}\left( 
    1 + \cos\left[ \min(\frac{r}{r_c}, 1) \pi \right] 
\right)
\end{equation} 
and $\cos\theta$ also has a vectorized form:
\begin{equation}
\cos\theta = \frac{\rijna^2 + \rikna^2 - \rjkna^2}{2 \rijna \cdot \rikna}
\end{equation}
The summations in $\mathbf{G}^{(2)}$ and $\mathbf{G}^{(4)}$ should go over all 
neighbors within $r_c$.

@@

The $3 \times 3$ virial stress tensor $\epsilon$ is an important metric to 
describe solids under deformation or external pressure. For arbitrary many-body 
interaction potentials, $\epsilon$ can be calculated with the following equation
\cite{lammps_stress}:
\begin{equation}
\label{eq:stress_lammps}
V \cdot \epsilon = \left(-\sum_{i=1}^{N}{\mathbf{r}_i^{(0)} \otimes f_i} - 
\sum_{\mathbf{n}}{\mathbf{h}^T\mathbf{n}} \otimes 
\sum_{i=1}^{N}{F^{\prime}_{i\mathbf{n}}}
\right)^T
\end{equation}
where $V$ is the volume, $\otimes$ indicates tensor product, $f_i$ is the total 
force acting on atom $i$ and $F^{\prime}_{i\mathbf{n}}$ is the partial force. 
Until now, most machine learning force-field models focus on fitting total 
energy and atomic forces, very few of them\cite{DeePMD,DeePMD_rl,DeePMD_kit} had 
attempted to include stress in their loss functions. 
The current expression of Equation \ref{eq:stress_lammps} is not very compatible 
with NN or other ML potentials since the partial forces are difficult to compute 
with vectorized operations. However, we find an "equivalent" form 
of Equation \ref{eq:stress_lammps}:
\begin{equation}
\label{eq:stress}
V \cdot \epsilon = -F^{T} R + 
\left(\frac{\partial E^{total}}{\partial \mathbf{h}}\right)^T \mathbf{h}
\end{equation}
where $R$ is the $N \times 3$ positions matrix and $F$ is the 
corresponding total forces matrix. This new form can be easily implemented 
within TensorFlow. The detailed derivation is given in the appendix.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.A
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{General design}
\label{section:principles}

Fig \ref{fig:program_design} shows the general design of TensorAlloy.
TensorAlloy has two phases: the training phase \textbf{(1)} and the prediction 
phase \textbf{(2)}. These two phases share many common requirements, but they 
also have their own concerns.

For the training phase, in order to make TensorAlloy a general-purpose and
efficient machine learning potential training program, these three key 
challenges shall be resolved:
\begin{itemize}

@@

The virtual-atom approach plays a central role in handling all the challenges 
above and unify the training and prediction phases. In the following sections, 
we will gradually discuss the details of this approach.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.1
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.58]{figures/Fig1.pdf}
\caption{\label{fig:program_design} The general design of TensorAlloy. 
\textbf{(1)} demonstrates the training phase while \textbf{(2)} represents the 
prediction phase.
}
\end{figure}

@@

Fig \ref{fig:Fig2} shows the preparation stage of VAP. The cutoff radius 
$r_{cut}$ must be fixed. This stage has two major steps. The first step is to 
build the global symbol list (GSL), which acts as the universal reference 
system. A virtual atom is inserted at the very first position of GSL. The 
benefits will be discussed later. To build GSL, all $\{\nelmax\}$ must be 
pre-determined where $\nelmax$ is the maximum appearances of element 
\textit{el} in any structure. $\nel$ represents the number of unique chemical 
elements in the dataset. Then, $\nvap$, the dimension of GSL, can be determined:
\begin{equation}
\label{eq:n_vap}
\nvap = 1 + \sum_{\mathrm{el}}{\nelmax}
\end{equation}
For the Ni-Mo dataset, $el \in [\mathrm{Ni}, \mathrm{Mo}]$ and 
$\nel = 2$. 
Table \ref{table:GSL} gives some examples of the GSL mapping. 'X' denotes the 
inserted virtual atom and Ni$_{108(54)}$ means the first 54 Ni atoms are mapped 
from the local stoichiometry. In the training phase, all $\nelmax$ are fixed.
In the prediction phase, $\nelmax$ shall be determined dynamically. One should 
note that the minimum value of $\nelmax$ should be 1 due to technical reason (so 
that the computation graph of TensorFlow can be properly executed). The codes to
get $\nelmax$ and GSL are given in Appendix(C). With GSL setup, atomic positions
and forces can be transformed as well. Appendix(D) gives the implementation of 
the virtual-atom mapping mechanism.

@@

Both the radial (Equation \ref{eq:g2}) and the angular (Equation \ref{eq:g4}) 
symmetry functions interact with interatomic distances only. In order to compute 
the descriptors, local atomic positions should be transformed to interatomic 
distances first. Fig \ref{fig:Fig3} demonstrates the pseudo graph of 
$\mathbf{R} \rightarrow \mathbf{r}^{\mathrm{GSL}}$ for radial symmetry function
descriptors:

@@

Now we can start the last step: computing radial symmetry descriptors 
$\mathbf{G}^{(2)}$ from $\mathbf{r}^{\mathrm{GSL}}$. For simplicity, we will use 
the prediction phase to demonstrate this final section.

@@

    \item[4.]
    In the training phase, $\nvap$ is a constant determined before training. In 
    the contrast, $\nvap$ should be a dynamic tensor 
    (or a \textit{tf.Placeholder} technically) in the prediction phase.
    Its value shall be runtimely determined according to the input structure. 
    Thus, the stoichiometry challenge mentioned in Section 
    \ref{section:principles} can be overcame.

@@

To train the model, we use the following root mean-squared error (RMSE) total 
loss function:
\begin{align}
\label{eq:loss}
\mathbf{Loss} & = \sqrt{\frac{1}{N_{b}}\sum_{i=1}^{N_{b}}{\left(
    E_{i} - E_{i}^{\mathrm{dft}}
\right)^2}} \nonumber \\
& + \chi_{\mathrm{f}}\sqrt{
    \frac{1}{3 N_{b} \nvap}\sum_{i}^{N_b}{\sum_{j}^{\nvap}{
        \sum_{\alpha}{
            \left(f_{ij\alpha} - f_{ij\alpha}^{\mathrm{dft}}\right)^2
        }
    }}
} \nonumber \\
& + \chi_{\mathrm{s}}\sqrt{\frac{1}{6N_b}\sum_{i}^{N_b}{
    \sum_{j}^{6}{
        \left(
            \epsilon^{\mathrm{voigt}}_{j} - \epsilon^{\mathrm{voigt,dft}}_{j}
        \right)^2
    }
}}
\end{align}
where $N_b$ is the mini-batch size and $\chi_{\mathrm{f}}$ and 
$\chi_{\mathrm{s}}$ are weights of force and stress losses. 
Stress tensors are converted to Voigt vectors. All energies are in \textbf{eV}, 
all forces are in \textbf{eV/\AA} and all stress components are in 
\textbf{eV/\AA}$\mathbf{^3}$. In most cases, we set $N_b$ to 50 or 100 and we
use the ADAM\cite{adam} optimizer with exponentially-decayed learning rate to 
minimize the loss function. The default activation function is softplus:
\begin{equation}
\sigma(x) = \log(1 + e^x)
\end{equation}
The kernel weights of all NNs are initialized with the 
Xvaier\cite{pmlr-v9-glorot10a} method and kernel biases are initialized with 
zeros.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 4. Discussions
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Discussions}
\label{section:discussions}

@@

    \item Ni: this model uses 400 Ni structures for training and 61 for 
    evaluation. The loss function only includes energy and force contributions.

    \item Mo: this model uses 250 Mo structures for training and 34 for 
    evaluation. The loss function includes all three (energy, force, stress) 
    types of contributions.

    \item Ni-Mo: this model uses 3673 structures for training and 300 for 
    evaluation. The loss function includes all three types of contributions.

@@

For all these models, the cutoff radius is set to 6.5 \AA. Only radial symmetry 
functions are used to compute atomic descriptors. The selected $\eta$ are 
0.1, 0.5, 1, 2, 4, 8, 12, 16, 20, 40 and $\omega$ are 0.0, 3.0. The hidden layer
sizes are 128, 64, 32. The activation function is softplus.
The learning rate starts from 0.01 and it will decay exponentially with rate 
0.95 for every 3000 steps.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table. 3
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{cccc|ccc|cc}
\hline
 & \multicolumn{3}{c|}{Energy (meV/atom)} & \multicolumn{3}{c|}{Force (eV/\AA)}
 & \multicolumn{2}{c}{Stress (GPa)} \\
\hline
Model & Mo & Ni & Ni-Mo & Mo & Ni & Ni-Mo & Mo & Ni-Mo \\
\hline
Mo & 4.9 (13.2) & & 16.5 (16.2) & 0.19 (0.25) & & 0.28 (0.29) & 0.24 (0.87) 
   & 0.75 \\
Ni & & 1.2 (1.2) & 4.5 (7.9) & & 0.04 (0.05) & 0.07 (0.11) & & 0.83 \\
Ni$_4$Mo & & & 4.1 (4.0) & & & 0.09 (0.14) & & 0.98 \\ 
Ni$_3$Mo & & & 4.5 (5.2) & & & 0.11 (0.16) & & 1.19 \\
Ni$_{\mathrm{Mo}}$ & & & 12.9 (22.7) & & & 0.09 (0.13) & & 0.34 \\
Mo$_{\mathrm{Ni}}$ & & & 12.9 (33.9) & & & 0.12 (0.55) & & 0.45 \\
Overall & & & 10.8 (22.5) & & & 0.11 (0.23) & & 0.59 \\
\hline
\end{tabular}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.7
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.64]{figures/Prediction-speed.pdf}
\caption{\label{fig:prediction_speed} Benchmarks of the Ni-Mo model in 
the prediction phase. The test structures are (MoNi)$_{x}$ where 
$x$ = $10^3, 15^3, 20^3, 25^3, 30^3, 35^3, 40^3$. The left figure \textbf{a)} 
demonstrates the overall time and the right figure \textbf{b)} compares the 
'Graph' time with/out GPU. Blue bars indicate the elapsed time (in seconds) for 
finding neighbors, orange bars denote the time for initializing VAP-GSL arrays 
and green bars and the numbers above represents the graph execution time (used 
for obtaining energy, forces and stress).
}
\end{figure}

The speed of the prediction phase is also a crucial metric for measuring 
interaction potentials. Fig \ref{fig:prediction_speed} summarizes the time 
distribution of the Ni-Mo model in the prediction phase on the GPU workstation. 
Here 'prediction' is defined as calculating energy, forces and stress tensor of 
an arbitrary \textit{ase.Atoms} object. In these tests, the primitive structure
$\mathrm{MoNi}$ is obtained by replacing one of the two bcc Mo atoms with a Ni 
atom. The test subjects $(\mathrm{MoNi}))_{x}$ are just supercells of the base 
structure. We choose $x \in [10^3, 15^3, 20^3, 25^3, 30^3, 35^3, 40^3]$ as
a typical molecular dynamics simulation for studying alloy properties needs at 
least thousands of atoms. The elaspsed time is splitted to three parts: 
'Neighbor' (blue) indicates the execution time of the \textit{neighbor\_list} 
routine of ASE for finding local neighbors (\textbf{(1)} of Fig \ref{fig:Fig3}),
'VAP' (orange) denotes the time to construct VAP-GSL arrays 
(\textbf{(2)} of Fig \ref{fig:Fig3}) and 'Graph' (green bars and the numbers 
above) represents the time of getting energy, forces and stress tensor by 
executing the computation graph. The smallest subject (2000 atoms) needs 1.6 
seconds and the largest system (128000 atoms) requires ~71.6 seconds on the 
workstation. Most of the time are spent on the Python-based single-thread 
'Neighbor' and 'VAP' routines \textemdash which may be greatly optimized by 
utilizing efficient C/C++/Fortran codes (e.g. Lammps). For the 'Graph' part, the 
execution time can be as small as 1 second (with GPU) or 4 seconds (without GPU) 
for the 128000-atoms structure. According to these tests, it's fairly practical 
to use TensorAlloy-trained symmetry function interaction potential on studying 
large alloy systems.

>>>>

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can provide density functional theory level predictions 
within negligible time. Currently, the symmetry function descriptor based 
atomistic neural network is the most widely used model for modeling alloys. 
To precisely describe complex potential energy surfaces, integrating advanced 
metrics, such as force or virial stress, into training can be of great help.  
In this work, we propose a virtual-atom approach to model the total energy of 
symmetry function descriptors based atomistic neural network. Our approach 
starts from atomic positions directly so that the derivations of forces and 
virial can be handled by TensorFlow automatically and efficiently. This new 
approach is implemented in our program TensorAlloy, which supports constructing 
machine learning interaction potentials for both \textit{molecules} and 
\textit{solids}. The QM7 and SNAP/Ni-Mo datasets are used to demostrate the 
performances of our program.
\end{abstract}

@@

In this work, we propose a new TensorFlow-dependent apporach (the virtual atom 
approach, VAP) to model the total energy expression of the symmetry function 
descriptor based neural network. 
Our apporach can move the entry of the computation graph from atomic descriptors 
back to the raw atomic positions. Thus, calculating the derivatives of total 
energy with respect to atomic positons (force) or lattice tensor (virial) can be 
handled by TensorFlow directly and automatically.

This paper is organized as follows. Section \ref{section:theory} introduces the 
theory of the symmetry function descriptor and the ANN framework. Section 
\ref{section:method} describes the virtual atom approach and the modified ANN 
model used by TensorAlloy. Section \ref{section:discussions} demostrates 
the performances of TensorAlloy on two public datasets. The appendix 
\ref{section:appendix} gives derivation and implementation details of key 
routines of VAP. The full implementation and demo codes can be obtained from 
GitHub (https://github.com/Bismarrck/vap) freely.

@@

There are many choices of $F_G$. The symmetry function descriptor, proposed 
by Parinello and Behler in 2007\cite{Behler}, is currently the most widely used 
atomic descriptor. The symmetry function descriptor consists of two types of 
functions: the radial symmetry function $\mathbf{G}^{(2)}$ and the angular 
symmetry function $\mathbf{G}^{(4)}$:
\begin{align}
\label{eq:G2}
\mathbf{G}^{(2)}_i(\eta, \omega) = & \sum_{j \neq i}{
    \sum_{\mathbf{n}}{g_2 (\eta, i, j, \omega)}
} \\
\label{eq:G4}
\mathbf{G}^{(4)}_i(\beta, \gamma, \zeta) 
= & 2^{1-\zeta}\sum_{j, k\neq j, k\neq i}{
    \sum_{\mathbf{n}_1}{\sum_{\mathbf{n}_2}{\sum_{\mathbf{n}_3}{
        g_4(\beta, \gamma, \zeta, 
          i, j, k, 
          \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3)
    }}}
} \\
\label{eq:g2}
g_2(\eta, i, j, \omega) = 
& \exp\left\{-\frac{\eta(\rijn - \omega)^2}{r_c^2} \right\} \cdot f_c(\rijn) \\
\label{eq:g4}
g_4(
    \beta, \gamma, \zeta, 
    i, j, k, 
    \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3) = & 
\left( 1 + \gamma\cos\theta \right)^\zeta 
    \exp \left\{ -\frac{\beta(\rijna^2 + \rikna^2 + \rjkna^2)}{r_c^2} \right\}
\nonumber \\
& \times f_c(\rijna)f_c(\rikna)f_c(\rjkna)
\end{align}
where $\eta$, $\omega$, $\beta$, $\gamma$ and $\zeta$ are empirically chosen 
parameters, $f_c(r)$ is a cutoff (damping) function and its original form is:
\begin{equation}
\label{eq:fc_origin}
f_c(r) =
    \begin{cases}
    0 & \text{$r > r_c$} \\
    \frac{1}{2} + \frac{1}{2}\cos\left(r / r_c \cdot \pi \right) & 
    \text{$r \leq r_c$}
    \end{cases}
\end{equation}
and $r_c$ is the cutoff radius. Equation \eqref{eq:fc_origin} can be further 
transformed to a vectorized expression:
\begin{equation}
\label{eq:fc}
f_c(r) = \frac{1}{2}\left( 
    1 + \cos\left[ \min(\frac{r}{r_c}, 1) \pi \right] 
\right)
\end{equation} 
and $\cos\theta$ also has a vectorized form:
\begin{equation}
\cos\theta = \frac{\rijna^2 + \rikna^2 - \rjkna^2}{2 \rijna \cdot \rikna}
\end{equation}
The summations in Equation \ref{eq:G2} and \ref{eq:G4} should go over all 
neighbors within $r_c$. For an atom $i$, $\mathbf{G}_i$ is a vector. For the 
entire structure, its overall $\mathbf{G}$ is a matrix with row $i$ represents 
the atomic descriptors of atom $i$ and column $j$ corresponds to a specifc 
$(\eta, \omega)$ pair or $(\beta, \gamma, \zeta)$ combination.

@@

The $3 \times 3$ virial tensor $\epsilon$ is an important metric to 
describe solids under deformation or external pressure. For arbitrary many-body 
interaction potentials, $\epsilon$ can be calculated with the following equation
\cite{lammps_stress}:
\begin{equation}
\label{eq:stress_lammps}
\epsilon = \left(-\sum_{i=1}^{N}{\mathbf{r}_i^{(0)} \otimes f_i} - 
\sum_{\mathbf{n}}{\mathbf{h}^T\mathbf{n}} \otimes 
\sum_{i=1}^{N}{F^{\prime}_{i\mathbf{n}}}
\right)^T
\end{equation}
where $\otimes$ indicates tensor product, $f_i$ is the total 
force acting on atom $i$ and $F^{\prime}_{i\mathbf{n}}$ is the partial force. 
Until now, most machine learning force-field models focus on fitting total 
energy and atomic forces, very few of them\cite{DeePMD,DeePMD_rl,DeePMD_kit} had 
attempted to include stress in their loss functions. 
The current expression of Equation \ref{eq:stress_lammps} is not very compatible 
with NN or other ML potentials since the partial forces are difficult to compute 
with vectorized operations. However, we find an "equivalent" form 
of Equation \ref{eq:stress_lammps}:
\begin{equation}
\label{eq:stress}
\epsilon = -F^{T} R + 
\left(\frac{\partial E^{total}}{\partial \mathbf{h}}\right)^T \mathbf{h}
\end{equation}
where $R$ is the $N \times 3$ positions matrix and $F$ is the 
corresponding total forces matrix. This new form can be easily implemented 
within TensorFlow. The detailed derivation is given in the appendix(A).

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.A
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Overview}
\label{section:overview}

Fig \ref{fig:program_design} shows the overview of TensorAlloy and the 
virtual-atom approach.
TensorAlloy has two phases: the training phase \textbf{(1)} and the prediction 
phase \textbf{(2)}. These two phases share many common requirements, but they 
also have their own concerns.

For the training phase, in order to make TensorAlloy a general-purpose and
efficient machine learning potential training program, there are some key 
technical requirements:
\begin{itemize}

@@

VAP plays a central role in handling all the technical challenge as it 
\textbf{\textit{provides a way to describe structures of various sizes and 
stoichiometries in unified and vectorized forms}}. 
In the following sections, we will gradually discuss the details of this 
approach.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.1
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.68]{figures/Fig1.pdf}
\caption{\label{fig:program_design} The general design of TensorAlloy. 
\textbf{(1)} demonstrates the training phase while \textbf{(2)} represents the 
prediction phase.
}
\end{figure}

@@

Fig \ref{fig:Fig2} shows the preparation stage of VAP. The cutoff radius 
$r_{cut}$ must be fixed. This stage has two major steps. The first step is to 
build the global symbol list (GSL), which acts as the universal reference 
system. A virtual atom is inserted at the very first position of GSL. 
To build GSL, all $\{\nelmax\}$ must be pre-determined where $\nelmax$ is the 
maximum appearances of element \textit{el} in any structure. $\nel$ represents 
the number of unique chemical elements in the dataset. Then, $\nvap$, the 
dimension of GSL, can be determined:
\begin{equation}
\label{eq:n_vap}
\nvap = 1 + \sum_{\mathrm{el}}{\nelmax}
\end{equation}
For the Ni-Mo dataset, $el \in [\mathrm{Ni}, \mathrm{Mo}]$ and 
$\nel = 2$. 
Table \ref{table:GSL} gives some examples of the GSL mapping. 'X' denotes the 
inserted virtual atom and Ni$_{108(54)}$ means the first 54 Ni atoms are mapped 
from the local stoichiometry. In the training phase, all $\nelmax$ are fixed.
In the prediction phase, $\nelmax$ shall be determined dynamically. One should 
note that the minimum value of $\nelmax$ should be 1 due to technical reason (so 
that the computation graph of TensorFlow can be properly executed). The codes to
get $\nelmax$ and GSL are given in Appendix(C). With GSL setup, atomic positions
and forces can be transformed as well. Appendix(D) gives the implementation of 
the virtual-atom mapping mechanism.

@@

Both the radial (Equation \ref{eq:g2}) and the angular (Equation \ref{eq:g4}) 
symmetry functions depends on interatomic distances only. In order to compute 
the descriptors, local atomic positions should be transformed to interatomic 
distances first. Fig \ref{fig:Fig3} demonstrates the pseudo graph of 
$\mathbf{R} \rightarrow \mathbf{r}^{\mathrm{GSL}}$ for radial symmetry function
descriptors:

@@

Now we can start computing radial symmetry descriptors 
$\mathbf{G}^{(2)}$ from $\mathbf{r}^{\mathrm{GSL}}$. For simplicity, we will use 
the prediction phase to demonstrate this final section.

@@

    \item[4.]
    In the training phase, $\nvap$ is a constant determined before training. In 
    the contrast, $\nvap$ should be a dynamic tensor 
    (or a \textit{tf.Placeholder} technically) in the prediction phase.
    Its value shall be runtimely determined according to the input structure.

@@

To train the model, we use the following root mean-squared error (RMSE) total 
loss function:
\begin{align}
\label{eq:loss}
\mathbf{Loss} & = \sqrt{\frac{1}{N_{b}}\sum_{i=1}^{N_{b}}{\left(
    E_{i} - E_{i}^{\mathrm{dft}}
\right)^2}} \nonumber \\
& + \chi_{\mathrm{f}}\sqrt{
    \frac{1}{3 N_{b} \nvap}\sum_{i}^{N_b}{\sum_{j}^{\nvap}{
        \sum_{\alpha}{
            \left(f_{ij\alpha} - f_{ij\alpha}^{\mathrm{dft}}\right)^2
        }
    }}
} \nonumber \\
& + \chi_{\mathrm{s}}\sqrt{\frac{1}{6N_b}\sum_{i}^{N_b}{
    \sum_{j}^{6}{
        \left(
            \epsilon^{\mathrm{voigt}}_{j} - \epsilon^{\mathrm{voigt,dft}}_{j}
        \right)^2
    }
}}
\end{align}
where $N_b$ is the mini-batch size and $\chi_{\mathrm{f}}$ and 
$\chi_{\mathrm{s}}$ are weights of force and virial losses. 
In most cases, we set $N_b$ to 50 or 100 and we
use the ADAM\cite{adam} optimizer with exponentially-decayed learning rate to 
minimize the loss function. The default activation function is softplus:
\begin{equation}
\sigma(x) = \log(1 + e^x)
\end{equation}
Softplus is a smooth approximation of the ReLU activation function.
The kernel weights of all NNs are initialized with the 
Xvaier method\cite{pmlr-v9-glorot10a} and kernel biases are initialized with 
zeros.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 4. Results and Discussions
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Results and Discussions}
\label{section:discussions}

@@

    \item \textbf{NN}(Ni): this model uses 400 Ni structures for training and 61 
    for evaluation. The loss function only includes energy and force 
    contributions.

    \item \textbf{NN}(Mo): this model uses 250 Mo structures for training and 34 
    for evaluation. The loss function includes all three (energy, force, stress) 
    types of contributions.

    \item \textbf{NN}(Mo-Ni): this model uses 3673 structures for training and 
    300 for evaluation. The loss function includes all three types of 
    contributions.

@@

The cutoff radius is set to 6.5 \AA. Only radial symmetry 
functions are used to compute atomic descriptors. The selected $\eta$ are 
0.1, 0.5, 1, 2, 4, 8, 12, 16, 20, 40 and $\omega$ are 0.0, 3.0. The hidden layer
sizes are 128, 64, 32. The activation function is softplus.
The learning rate starts from 0.01 and it will decay exponentially with rate 
0.95 for every 3000 steps.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table. 3
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{cccc|ccc|cc}
\hline
 & \multicolumn{3}{c|}{Energy (meV/atom)} & \multicolumn{3}{c|}{Force (eV/\AA)}
 & \multicolumn{2}{c}{Stress (GPa)} \\
\hline
NN Model & Mo & Ni & Ni-Mo & Mo & Ni & Ni-Mo & Mo & Ni-Mo \\
\hline
Mo & 4.9 (13.2) & & 16.5 (16.2) & 0.19 (0.25) & & 0.28 (0.29) & 0.24 (0.87) 
   & 0.75 \\
Ni & & 1.2 (1.2) & 4.5 (7.9) & & 0.04 (0.05) & 0.07 (0.11) & & 0.83 \\
Ni$_4$Mo & & & 4.1 (4.0) & & & 0.09 (0.14) & & 0.98 \\ 
Ni$_3$Mo & & & 4.5 (5.2) & & & 0.11 (0.16) & & 1.19 \\
Ni$_{\mathrm{Mo}}$ & & & 12.9 (22.7) & & & 0.09 (0.13) & & 0.34 \\
Mo$_{\mathrm{Ni}}$ & & & 12.9 (33.9) & & & 0.12 (0.55) & & 0.45 \\
Overall & & & 10.8 (22.5) & & & 0.11 (0.23) & & 0.59 \\
\hline
\end{tabular}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.7
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.64]{figures/Prediction-speed.pdf}
\caption{\label{fig:prediction_speed} Benchmarks of the \textbf{NN}(Mo-Ni) model 
in the prediction phase. The test structures are (MoNi)$_{x}$ where 
$x$ = $10^3, 15^3, 20^3, 25^3, 30^3, 35^3, 40^3$. The left figure \textbf{a)} 
demonstrates the overall time and the right figure \textbf{b)} compares the 
'Graph' time with/out GPU. Blue bars indicate the elapsed time (in seconds) for 
finding neighbors, orange bars denote the time for initializing VAP-GSL arrays 
and green bars and the numbers above represents the graph execution time (used 
for obtaining energy, forces and stress).
}
\end{figure}

The speed of the prediction phase is also a crucial metric for measuring 
interaction potentials. Fig \ref{fig:prediction_speed} summarizes the time 
distribution of the \textbf{NN}(Ni-Mo) model in the prediction phase on the GPU 
workstation. 
Here 'prediction' is defined as calculating energy, forces and stress tensor of 
an arbitrary \textit{ase.Atoms} object. In these tests, the primitive structure
$\mathrm{MoNi}$ is obtained by replacing one of the two bcc Mo atoms with a Ni 
atom. The test subjects $(\mathrm{MoNi}))_{x}$ are just supercells of the base 
structure. We choose $x \in [10^3, 15^3, 20^3, 25^3, 30^3, 35^3, 40^3]$ as
a typical molecular dynamics simulation for studying alloy properties needs at 
least thousands of atoms. The elaspsed time is splitted to three parts: 
'Neighbor' (blue) indicates the execution time of the \textit{neighbor\_list} 
routine of ASE for finding local neighbors (\textbf{(1)} of Fig \ref{fig:Fig3}),
'VAP' (orange) denotes the time to construct VAP-GSL arrays 
(\textbf{(2)} of Fig \ref{fig:Fig3}) and 'Graph' (green bars and the numbers 
above) represents the time of getting energy, forces and stress tensor by 
executing the computation graph. The smallest subject (2000 atoms) needs 1.6 
seconds and the largest system (128000 atoms) requires ~71.6 seconds on the 
workstation. Most of the time are spent on the Python-based single-thread 
'Neighbor' and 'VAP' routines \textemdash which may be greatly optimized by 
utilizing efficient C/C++/Fortran codes (e.g. Lammps). For the 'Graph' part, the 
execution time can be as small as 1 second (with GPU) or 4 seconds (without GPU) 
for the 128000-atoms structure. According to these tests, it's fairly practical 
to use TensorAlloy-trained symmetry function interaction potential on studying 
large alloy systems.
